Determinism, Complexity, and Predictability in Computer Performance

### 2014-08-01 results.tex
added mention of DVS fits

Figure~\ref{fig:wpe_vs_mase_best} plots the WPE values versus the
corresponding MASE values of the \emph{best} prediction for each of
the 120 time series in this study.
\begin{figure}
  \centering
  \includegraphics[width=1.1\columnwidth]{figs/new_prediction_vs_entropy}
  \caption{Weighted permutation entropy vs. mean absolute scaled error
    (MASE) of the best prediction of each time series.  The solid
    curve is a least-squares log fit of these points.
%
% of all the points except those from {\tt dgesdd$_1$}, which we have
% excluded for reasons explained in the text.
%
The dashed curves reflect the standard deviation of the model in its
parameter space.  The meaning of the shaded region is explained in the
text.}
  \label{fig:wpe_vs_mase_best}
\end{figure}
%
% As the plot shows, the best-case prediction error for each trace from
% these eight systems is roughly proportional to the log of the weighted
% permutation entropy, which is consistent with our first conjecture.
%
There is an obvious upward trend, which is consistent with the notion
that there is a pattern in the WPE-MASE relationship.  However, a
simple linear fit is a bad idea here.  First, any signal with zero
entropy should be perfectly predictable (i.e., MASE $\approx 0$), so
any curve fitted to these data should pass through the origin.
Moreover, WPE does not grow without bound, so one would expect the
patterns in the WPE-MASE pairs to reach some sort of asymptote.  For
these reasons, we chose to fit a function of the form $y = a \log(b x
+ 1)$ to these points, with $y =$ WPE and $x=$ MASE\footnote{The
  specific values of the coefficiencts are $a=7.97 \times 10^{-2}$ and
  $b=1.52 \times 10^3$.}.  The solid curve in the figure shows this
fit; the dashed curves show the standard deviation of this model in
its parameter space: i.e., $y = a \log(b x + 1)$ with $\pm$ one
standard deviation on each of the two parameters.  Points that fall
within this deviation volume (light grey) correspond to predictions
that are comparable to the best ones found in this study; points that
fall \emph{above} that volume (dark grey) are better still.  We chose
to truncate the shaded region because of a subtle point regarding the
MASE of an ideal predictor, which should not be larger than 1 unless
the training and test signals are different.  This is discussed at
more length below.

@@

The second effect that can skew MASE scores is nonstationarity.  Since
this metric is normalized by the error of a random-walk forecast
\emph{on the training signal}, differences between the test signal and
training signal can create issues.  This is why the MASE values in
Table~\ref{tab:error} are not identically one for every random-walk
forecast of every time series: the last 10\% of these signals is
significantly different from the first 90\%.  The deviation from 1.00
will depend on the process---whether it has multiple regimes, what
those regimes look like, and how it switches between them---as well as
the experimental setup (e.g., sensor precision and data length).  For
the processes studied here, these effects do not cause the MASE values
to exceed 1.15, but pathological situations (e.g., a huge switch in
scale right at the training/test signal boundary, or a signal that
simply grows exponentially) could produce higher values.  This
suggests another potentially useful heuristic: if the MASE of a
random-walk prediction of a time series is significantly different
from 1, this could be an indication that the signal is nonstationary.
We are in the process of exploring this idea.

>>>>

Figure~\ref{fig:wpe_vs_mase_best} plots the WPE values versus the
corresponding MASE values of the \emph{best} prediction for each of
the 120 time series in this study.
\begin{figure}
  \centering
  \includegraphics[width=1.1\columnwidth]{figs/new_prediction_vs_entropy}
  \caption{Weighted permutation entropy vs. mean absolute scaled error
    (MASE) of the best prediction of each time series.  The solid
    curve is a least-squares log fit of these points.
%
% of all the points except those from {\tt dgesdd$_1$}, which we have
% excluded for reasons explained in the text.
%
The dashed curves reflect the standard deviation of the model in its
parameter space.  The meaning of the shaded region is explained in the
text. \alert{Do you think we should explain that here, rather than
  punting to the text?  It is, after all, the primary claim of the
  paper...}}
  \label{fig:wpe_vs_mase_best}
\end{figure}
%
% As the plot shows, the best-case prediction error for each trace from
% these eight systems is roughly proportional to the log of the weighted
% permutation entropy, which is consistent with our first conjecture.
%
There is an obvious upward trend, which is consistent with the notion
that there is a pattern in the WPE-MASE relationship.  However, a
simple linear fit is a bad idea here.  First, any signal with zero
entropy should be perfectly predictable (i.e., MASE $\approx 0$), so
any curve fitted to these data should pass through the origin.
Moreover, WPE does not grow without bound, so one would expect the
patterns in the WPE-MASE pairs to reach some sort of asymptote.  For
these reasons, we chose to fit a function of the form $y = a \log(b x
+ 1)$ to these points, with $y =$ WPE and $x=$ MASE\footnote{The
  specific values of the coefficiencts are $a=7.97 \times 10^{-2}$ and
  $b=1.52 \times 10^3$.}.  The solid curve in the figure shows this
fit; the dashed curves show the standard deviation of this model in
its parameter space: i.e., $y = a \log(b x + 1)$ with $\pm$ one
standard deviation on each of the two parameters.  Points that fall
within this deviation volume (light grey) correspond to predictions
that are comparable to the best ones found in this study; points that
fall \emph{above} that volume (dark grey) are better still.  We chose
to truncate the shaded region because of a subtle point regarding the
MASE of an ideal predictor, which should not be larger than 1 unless
the training and test signals are different.  This is discussed at
more length below.

@@

The second effect that can skew MASE scores is nonstationarity.  Since
this metric is normalized by the error of a random-walk forecast
\emph{on the training signal}, differences between the test signal and
training signal can create issues.  This is why the MASE values in
Table~\ref{tab:error} are not identically one for every random-walk
forecast of every time series: the last 10\% of these signals is
significantly different from the first 90\%.  The deviation from 1.00
will depend on the process---whether it has multiple regimes, what
those regimes look like, and how it switches between them---as well as
the experimental setup (e.g., sensor precision and data length).  For
the processes studied here, these effects do not cause the MASE values
to exceed 1.15, but pathological situations (e.g., a huge switch in
scale right at the training/test signal boundary, or a signal that
simply grows exponentially) could produce higher values.  This
suggests another potentially useful heuristic: if the MASE of a
random-walk prediction of a time series is significantly different
from 1, this could be an indication that the signal is nonstationary.
We are in the process of exploring this idea and its relationship to
the DVS fits mentioned in Section~\ref{sec:related}.
